### 2014-09-30 doc/thesis/system.tex
Trying to start writing about the type system

\section{Type Language}

\begin{figure}[!ht]
\textbf{Type Language}\\
\dstart
$$
\begin{array}{rlr}
\multicolumn{3}{c}{\textbf{First-level types}}\\
T ::= & \;\; L & \textit{literal types}\\
& | \; B & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; T \cup T & \textit{disjoint union types}\\
& | \; P \rightarrow R & \textit{function types}\\
& | \; \{ \}_{c|o|u} & \textit{empty table type}\\
& | \; \{F, ..., F\}_{c|o|u} & \textit{table types}\\
& | \; X & \textit{type variables}\\
& | \; \mu X.T & \textit{recursive types}\\
& | \; X_{i} & \textit{projection types}\\
L ::= & \False \; | \; \True \; | \; {\it number} \; | \; {\it string} & \\
B ::= & \Boolean \; | \; \Number \; | \; \String & \\
F ::= & K:T \; | \; \Const \; K:T & \textit{field types}\\ 
K ::= & L \; | \; B \; | \; \Value \; | \; \Any & \textit{key types}\\
\multicolumn{3}{c}{}\\
\multicolumn{3}{c}{\textbf{Second-level types}}\\
S ::= & \;\; T &\\
& | \; T* & \textit{vararg types}\\
& | \; T \times S & \textit{tuple types}\\
P ::= & \Void \; | \; S & \textit{parameters list type}\\
R ::= & \Void \; | \; S \; | \; S \cup R & \textit{return types}\\
\end{array}
$$
\dend
\caption{The abstract syntax of Typed Lua types}
\label{fig:typelang}
\end{figure}

Figure \ref{fig:typelang} presents the abstract syntax of the
Typed Lua types.
We split Typed Lua types into two categories:
\emph{first-level types} and \emph{second-level types}.
We use first-level types to represent Lua values and
second-level types to type expression lists, multiple assignments,
and function applications.

\section{Subtyping}

@@

\section{Consistent-Subtyping}

@@

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rl}
s ::= & \;\; \mathbf{skip} \; | \;
s_{1} \; ; \; s_{2} \; | \;
\vec{l} = \vec{e}  \; | \;
\mathbf{while} \; e \; \mathbf{do} \; s \; | \;
\mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}\\
& | \; \mathbf{local} \; \vec{n}{:}\vec{T} = \vec{e} \; \mathbf{in} \; s \; | \;
\mathbf{rec} \; n{:}T = f \; \mathbf{in} \; s \; | \;
\mathbf{return} \; \vec{e} \; | \;
a\\
e ::= & \;\; \mathbf{nil} \; | \;
\mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it number} \; | \;
{\it string} \; | \;
{...} \; | \;
f \; | \;
\{ \; \vec{c} \; \} \\
& | \; e_{1} + e_{2} \; | \;
e_{1} \; {..} \; e_{2} \; | \;
e_{1} == e_{2} \; | \;
e_{1} < e_{2} \; | \;
e_{1} \; \mathbf{and} \; e_{2} \; | \;
e_{1} \; \mathbf{or} \; e_{2} \\
& | \; \mathbf{not} \; e \; | \;
- e \; | \;
\# \; e \; | \;
a \; | \;
l \; | \;
(e) \; | \;
{<}t{>} \; e\\
l ::= & \; n \; | \;
e_{1}[e_{2}]\\
a ::= & \; e_{1}(\vec{e_{2}}) \; | \;
e_{1}{:}n(\vec{e_{2}})\\
f ::= & \; \mathbf{fun} \; (){:}R \; s \; | \;
\mathbf{fun} \; ({...}{:}T){:}R \; s \; | \;
\mathbf{fun} \; (\vec{n}{:}\vec{T}){:}R \; s \; | \;
\mathbf{fun} \; (\vec{n}{:}\vec{T},{...}{:}T){:}R \; s\\
c ::= & \; [e_{1}] = e_{2}\\
n ::= & \; {\it name}\\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}

>>>>

In the previous chapter we presented an informal overview of Typed Lua.
We showed that programmers can use Typed Lua to combine static and dynamic
typing in the same code.

In this chapter we present the formalization of Typed Lua.
We use typing rules to show how we implemented the features that we
discussed in the previous chapter.

\section{Types}

\begin{figure}[!ht]
\textbf{Type Language}\\
\dstart
$$
\begin{array}{rlr}
\multicolumn{3}{c}{\textbf{First-level types}}\\
t ::= & \;\; l & \textit{literal types}\\
& | \; b & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; t \cup t & \textit{disjoint union types}\\
& | \; p \rightarrow r & \textit{function types}\\
& | \; \{f_{i}, ..., f_{n}\}_{c|o|u} & \textit{table types}\\
& | \; x & \textit{type variables}\\
& | \; \mu x.t & \textit{recursive types}\\
& | \; x_{i} & \textit{projection types}\\
l ::= & \False \; | \; \True \; | \; {\it number} \; | \; {\it string} & \\
b ::= & \Boolean \; | \; \Number \; | \; \String & \\
f ::= & k:t \; | \; \Const \; k:t & \textit{field types}\\ 
k ::= & l \; | \; b \; | \; \Any & \textit{key types}\\
\multicolumn{3}{c}{}\\
\multicolumn{3}{c}{\textbf{Second-level types}}\\
s ::= & \;\; t &\\
& | \; t* & \textit{variadic types}\\
& | \; t \times s & \textit{tuple types}\\
p ::= & \Void \; | \; s & \textit{parameters list type}\\
r ::= & \Void \; | \; s \; | \; s \cup r & \textit{return types}\\
\end{array}
$$
\dend
\caption{The abstract syntax of Typed Lua types}
\label{fig:typelang}
\end{figure}

Figure \ref{fig:typelang} presents the abstract syntax of the
Typed Lua types.
Typed Lua splits types into two categories:
\emph{first-level types} and \emph{second-level types}.
Typed Lua uses first-level types to represent Lua values and
second-level types to represent expression lists, and it also
uses them to type multiple assignments and function applications.
First-level types include literal types, base types, the type $\Nil$,
the top type $\Value$, the dynamic type $\Any$, the type $\Self$,
union types, function types, table types, recursive types, and
projection types.
Second-level types include the type $\Void$, tuples of first-level
types optionally ending in a variadic type, and union of these tuples.

@@

@@

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rl}
s ::= & \;\; \mathbf{skip} \; | \;
s_{1} \; ; \; s_{2} \; | \;
\vec{l} = \vec{e}  \; | \;
\mathbf{while} \; e \; \mathbf{do} \; s \; | \;
\mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}\\
& | \; \mathbf{local} \; \vec{n}{:}\vec{t} = \vec{e} \; \mathbf{in} \; s \; | \;
\mathbf{rec} \; n{:}t = f \; \mathbf{in} \; s \; | \;
\mathbf{return} \; \vec{e} \; | \;
a\\
e ::= & \;\; \mathbf{nil} \; | \;
\mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it number} \; | \;
{\it string} \; | \;
{...} \; | \;
f \; | \;
\{ \; \vec{c} \; \} \\
& | \; e_{1} + e_{2} \; | \;
e_{1} \; {..} \; e_{2} \; | \;
e_{1} == e_{2} \; | \;
e_{1} < e_{2} \; | \;
e_{1} \; \mathbf{and} \; e_{2} \; | \;
e_{1} \; \mathbf{or} \; e_{2} \\
& | \; \mathbf{not} \; e \; | \;
- e \; | \;
\# \; e \; | \;
a \; | \;
l \; | \;
(e) \; | \;
{<}t{>} \; e\\
l ::= & \; n \; | \;
e_{1}[e_{2}]\\
a ::= & \; e_{1}(\vec{e_{2}}) \; | \;
e_{1}{:}n(\vec{e_{2}})\\
f ::= & \; \mathbf{fun} \; (){:}r \; s \; | \;
\mathbf{fun} \; ({...}{:}t){:}r \; s \; | \;
\mathbf{fun} \; (\vec{n}{:}\vec{t}){:}r \; s \; | \;
\mathbf{fun} \; (\vec{n}{:}\vec{t},{...}{:}t){:}r \; s\\
c ::= & \; [e_{1}] = e_{2}\\
n ::= & \; {\it name}\\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}
